Live Functional Programming with Typed Holes

### 2018-11-10 intro.tex
tweaks to spacing
format tweaking after adding copyright information

% !TEX root = hazelnut-dynamics.tex
\vspace{-5px}
\newcommand{\introSec}{Introduction}
\section{\protect\introSec} % don't like the all-caps thing that the template does, so protecting it from that
\label{sec:intro}

@@

Sec.~\ref{sec:calculus} then abstracts away the inessential details of the language and user interface and makes the  intuitions developed in Sec.~\ref{sec:examples} formally precise by detailing the primary contribution of this paper: a core calculus, \HazelnutLive, that supports evaluating incomplete expressions and tracking hole closures. 
% The dynamic semantics of \HazelnutLive is a small-step reduction semantics equipped with clean type safety theorems (including, notably, a Progress theorem for incomplete expressions). 
Sec.~\ref{sec:agda-mechanization} outlines our Agda-based mechanization of \HazelnutLive, which is included in the supplemental material and is also available from the following URL:

@@

\noindent
Sec.~\ref{sec:implementation} states the continuity invariant, which formally solves the gap problem, as a corollary of the primary theorems of \Hazelnut and \HazelnutLive. It also provides some additional details on the implementation of \Hazel. A snapshot of the implementation is included in the supplemental material. An online version of the ongoing implementation of \Hazel is available from \url{hazel.org}, and the source code is available from the following URL:

>>>>

% !TEX root = hazelnut-dynamics.tex
\ifarxiv \clearpage \fi
\newcommand{\introSec}{Introduction}
\section{\introSec} 
\label{sec:intro}

@@

Sec.~\ref{sec:calculus} then abstracts away the inessential details of the language and user interface and makes the  intuitions developed in Sec.~\ref{sec:examples} formally precise by detailing the primary contribution of this paper: a core calculus, \HazelnutLive, that supports evaluating incomplete expressions and tracking hole closures. 
% The dynamic semantics of \HazelnutLive is a small-step reduction semantics equipped with clean type safety theorems (including, notably, a Progress theorem for incomplete expressions). 
Sec.~\ref{sec:agda-mechanization} outlines our Agda-based mechanization of \HazelnutLive, which is included in the archived artifact and is also available from the following URL:

@@

\noindent
Sec.~\ref{sec:implementation} states the continuity invariant, which formally solves the gap problem, as a corollary of the primary theorems of \Hazelnut and \HazelnutLive. It also provides some additional details on the implementation of \Hazel. A snapshot of the implementation is included in the archived artifact. An online version of the ongoing implementation of \Hazel is available from \url{hazel.org}, and the source code is available from the following URL:

### 2018-11-10 related-work.tex
make section titles all caps again (blah)
format tweaking after adding copyright information

% !TEX root = hazelnut-dynamics.tex
\vspace{-5px}
\newcommand{\relatedWorkSection}{Related and Future Work}
\section{\protect\relatedWorkSection} % don't like the all-caps thing that the template does, so protecting it from that
\label{sec:relatedWork}
\vspace{-2px}

@@

%
There is also more general work on explicit substitutions, which records all  substitutions, not just substitutions around metavariables, explicitly \cite{Abadi:1991fr,levy1999explicit,Abadi:1990ys}.  \citet{DBLP:journals/corr/abs-1009-2789} have developed a theory of explicit substitutions for CMTT, which, following other work on explicit substitutions and environmental abstract machines \cite{DBLP:journals/tcs/Curien91}, could be useful in implementing \HazelnutLive more efficiently by delaying both standard and contextual substitutions until needed during evaluation. We leave this and other questions of fast implementation (e.g. using thunks to encapsulate indeterminate sub-expressions) to future work.

@@

Expression holes also often appear in the context of program synthesis, serving as
placeholders in \emph{templates}~\cite{srivastava2013template} or
\emph{sketches}~\cite{solar2009sketching} to be filled in by an expression
synthesis engine. We leave to future work the exciting possibility of combining these approaches. For example, one can imagine running an incomplete program as described in this paper and then adding tests or assertions about the value that a hole should take on under each of the different associated hole closures, via the live context inspector. These could serve as  constraints for use by a type-and-example-driven synthesis engine \cite{DBLP:conf/popl/FrankleOWZ16}. Relatedly, \emph{prorogued programming} proposes soliciting an external source, e.g. the programmer, for a suitable value when a hole is encountered \cite{DBLP:conf/oopsla/AfshariBS12}.

>>>>

% !TEX root = hazelnut-dynamics.tex
\newcommand{\relatedWorkSection}{Related and Future Work}
\section{\relatedWorkSection} % don't like the all-caps thing that the template does, so protecting it from that
\label{sec:relatedWork}

@@

%
There is also more general work on explicit substitutions, which records all  substitutions, not just substitutions around metavariables, explicitly \cite{Abadi:1991fr,levy1999explicit}.  \citet{DBLP:journals/corr/abs-1009-2789} have developed a theory of explicit substitutions for CMTT, which, following other work on explicit substitutions and environmental abstract machines \cite{DBLP:journals/tcs/Curien91}, could be useful in implementing \HazelnutLive more efficiently by delaying both standard and contextual substitutions until needed during evaluation. We leave this and other questions of fast implementation (e.g. using thunks to encapsulate indeterminate sub-expressions) to future work.

@@

Expression holes also often appear in the context of program synthesis, serving as
placeholders in \emph{templates}~\cite{srivastava2013template} or
\emph{sketches}~\cite{solar2009sketching} to be filled in by an expression
synthesis engine. We leave to future work the exciting possibility of combining these approaches. For example, one can imagine running an incomplete program as described in this paper and then adding tests or assertions about the value that a hole should take on under each of the different associated hole closures, via the live context inspector. These could serve as  constraints for use by a type-and-example-driven synthesis engine \cite{DBLP:conf/popl/FrankleOWZ16}. Relatedly, \emph{prorogued programming} proposes soliciting an external source, e.g. the programmer, for a suitable value when a hole instance is encountered \cite{DBLP:conf/oopsla/AfshariBS12}.
